---
title: "Reinforcement Learning for Business"
author: "Lars Relund Nielsen"
output:
  xaringan::moon_reader:
    css: "./libs/slides.css"
    lib_dir: libs
    seal: false
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r, child = "setup.Rmd", echo=FALSE}
```

```{r include=FALSE, eval=FALSE}
# to preview
xaringan::inf_mr(cast_from = ".")
```

layout: true

---

## Reinforcement Learning for Business

- The purpose of this course is to give an introduction and knowledge about reinforcement learning (RL). 
- RL is an approach of modelling sequential decision making problems such as 
  - controlling an inventory (decision: wait, buy, amount) 
  - maintain a spare-part (decision: wait, maintain)
  - playing backgammon (decision: how to move the checkers)
- 10 ECTS.
- Lectures/tutorials each week.
- Exercises and a project are given to support your learning.
<!-- - R is used as programming tool.  -->
- 20 minutes oral exam based on a project.

<!-- For more information see the [course description](https://kursuskatalog.au.dk/en?search=Reinforcement%20Learning%20for%20Business). -->


---

## Reinforcement learning can be seen as 

* An approach for learning good decision making under uncertainty from experience.
* Finding a good sequence of decisions over time.
* A model where given a state of a system, the agent wants to take actions to maximize future reward. Often the agent does not know the underlying setting and, thus, is bound to learn from experience.
* A way of modelling intuition. Our goal as an agent is to find good decisions/actions that maximize the total reward. That is, learn the reward estimate of a given action.
* A mathematical models for learning-based decision making.
  - Given state $x$ you take an action and observe the reward $r$ and the new state $x'$.
  - Your goal is to find a policy that optimize the total reward function.

Note current decisions have an impact on the future. 





```{r links, child="../book/links.md"}
```

```{r postprocess, include=FALSE}
system2("Rscript", args = "-e 'rmarkdown::render(\"index.Rmd\", quiet = TRUE)'")
file.copy("./slides.css", "./libs/", overwrite = T)
```
