---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Introduction to the book {#mod-intro}

```{r, code = readLines("common.R"), cache = FALSE, include=FALSE}
```


```{r include = FALSE, cache=FALSE}
library(tidyverse)
library(devtools)
library(DT)
library(htmltools)
library(rmarkdown)
library(bsplus)
library(here)
#tagList(html_dependency_font_awesome())
```


This site/book contains course notes for the course "Reinforcement Learning for Business" held at [Aarhus BSS][BSS]. The book show the learning path for the course modules. The course is an elective course mainly for the [Operations and Supply Chain Analytics][osca-programme] and [Business Intelligence][bi-programme] programme and intended to give you an introduction to Reinforcement Learning (RL). You can expect the book to be updated when the course runs. The date listed above is the last time the book was updated.


__Learning path diagram__

Click/hover the nodes to follow links and see details.

```{r, echo=FALSE, out.width="100%", fig.asp=NA}
# gs4_deauth()
# # bring your own app via client id (aka key) and secret
# google_app <- httr::oauth_app(
#   "tfa",
#   key = "afss",
#   secret = "test"
# )
# google_key <- "YOUR_API_KEY"
# gs4_auth_configure(app = google_app, api_key = google_key)
# 
# # confirm the changes
# gs4_oauth_app()
# gs4_api_key()

g <- create_learning_path(
   url = "https://docs.google.com/spreadsheets/d/1bBe42LHK-bE7CsU9eNBzi_7VNjbmv-Ybr7285pE61jM/edit?usp=sharing", 
   sheet = "intro", 
   x_legend = 4.43, y_legend = 0,
   margin_node = 0.2)
render_graph(g, height = 300)
```

## Learning outcomes {#lo-intro}

By the end of this module, you are expected to:

* Memorize the purpose of the book.
* Describe what the term Business Analytics mean.
* Identify pros and cons of using Excel, VBA and R.
* Describe how a computer works.
* Describe what an algorithm is.
* Know how the book is organized.

The learning outcomes relate to the [overall learning goals](#lg-course) number 1, 3 and 5 of the course.

<!-- SOLO increasing: identify · memorise · name · do simple procedure · collect data · -->
<!-- enumerate · describe · interpret · formulate · list · paraphrase · combine · do -->
<!-- algorithms · compare · contrast · explain causes · analyse · relate · derive · -->
<!-- evaluate · apply · argue · theorise · generalise · hypothesise · solve · reflect -->

## Purpose for the book {#purpose}

Since the amount of available data has increased extensively in many companies, there is a need for analysts with the ability to do tasks within Analytics. For instance, extract relevant data and perform valid quantitative analysis. Clearly, it is also important that the analyst can communicate the results of the analysis to their surroundings. This requires for the analyst to be particularly qualified in handling IT based tools beyond e.g. basic Excel. 

Business Analytics (BA) (or just Analytics) refers to the scientific process of transforming data into insight for making better decisions in business. BA can both be seen as the complete decision making process for solving a business problem and the creation of business value by integration of concepts, methods and data. As a process, it can be characterized by descriptive, predictive, and prescriptive model building using data sources. For a full definition see the [appendix](#sec:ba).




## How this book is organized {#organization}

Module \@ref(mod-intro) (this module) gives a short introduction to the book. Next, the book consists of different parts each containing teaching modules about specific topics:

   * [Part I](#mod-vba-intro) gives you an introduction to VBA and how to program using VBA in Excel.

   <!-- * [Part II](#mod-vba-datatypes) gives you an overview over programming in VBA including datatypes (Module \@ref(mod-vba-datatypes)), procedures (Module \@ref(mod-vba-procedures)), loop and conditional statements (Module \@ref(mod-vba-loops-cond)) and the object model (Module \@ref(mod-vba-object-model)). -->
   <!-- * [Part III](#mod-vba-random-numbers) contains extra modules useful when coding in VBA.  -->
   <!-- * [Part IV](#mod-vba-project) presents the mandatory VBA project. -->
   <!-- * [Part V](#mod-r-install) focuses on helping you install the needed programs on your computer (Module \@ref(mod-r-install)) and gives you a short introduction to R (Module \@ref(mod-r-workflow)). -->
   <!-- * [Part VI](#mod-r-loops-cond) gives you an overview over programming in R including loops and conditionals (Module \@ref(mod-r-loops-cond)) and functions (Module \@ref(mod-r-functions)). -->

   <!-- * [Part VII](#mod-r-tidyverse-intro) the focus is on import/export, tidy and transformation of data. Module \@ref(mod-r-tidyverse-intro) first gives you an introduction to the [tidyverse][tidyverse-main-page] packages and introduces you to literal programming using R Markdown. Next, Module \@ref(mod-r-io) shows you how to export and import data. Finally, Module \@ref(mod-r-transform) focuses on transformation of data. -->
   <!-- * [Part VIII](#mod-r-plot) considers visualization of data in R (Module \@ref(mod-r-plot)). -->
   <!-- * [Part IX](#mod-project) presents the mandatory R project. -->
   <!-- * [Part X](#mod-r-maps) contains extra modules useful when coding in R.  -->

The appendix contains different modules that may be helpful for you including hints on [how to work in groups](#groups), [how to get help](#help) if you are stuck and [how to annotate](#annotate) the course notes.
   


## Acknowledgements {#ack} 

Some of the materials in this book are taken from various places

* The bookdown skeleton and some notes are based on the [Tools for Analytics][tfa-course] course.

<!-- * Some parts in Module \@ref(mod-intro) are inspired by [Chapter 1](https://r4ds.had.co.nz/introduction.html) in @r4ds. -->
<!-- * The VBA modules are inspired by the book @vba-book. This also holds for some of the exercises.  -->
<!-- * Module \@ref(mod-r-install) is inspired by [Chapter 1](https://stat545.com/block000_r-rstudio-install.html) in @stat545. -->
<!-- * Module \@ref(mod-r-workflow) is using some text and images from [Chapter 1](https://moderndive.netlify.app/1-getting-started.html) in @moderndrive and [Chapter 2](https://stat545.com/block002_hello-r-workspace-wd-project.html) in @stat545. A few exercises are inspired by [Chapter 2](https://rafalab.github.io/dsbook/r-basics.html) in @dsbook. -->
<!-- * Notes about git and GitHub in the appendix are based on @hgit. -->
<!-- * Exercise \@ref(ex-gapminder1) is a revision of Chapters 6-7 in @stat545. -->
<!-- * Exercise \@ref(ex-babynames1) is a revision of [Session 3](https://github.com/rstudio-education/welcome-to-the-tidyverse/blob/master/03-Transform/03-Transform-Exercises.Rmd) in the [Welcome to the tidyverse][course-welcome-to-the-tidyverse] course. -->
<!-- * Exercise \@ref(ex-gapminder2) is a revision of [Chapter 9](https://rafalab.github.io/dsbook/gapminder.html) in @dsbook. -->
<!-- * Exercise \@ref(ex-covid19) is inspired by the COVID19 application [exercise](https://raw.githubusercontent.com/rstudio-education/datascience-box/master/course-materials/application-exercises/ae-01b-covid/covid.Rmd) at the [data science in a box](https://github.com/rstudio-education/datascience-box) course. -->
<!-- * Exercise \@ref(ex-lego) is inspired by the Lego homework [exercise](https://github.com/rstudio-education/datascience-box/blob/master/course-materials/hw-instructions/hw-04/hw-04-legos-instructors.Rmd) at the [data science in a box](https://github.com/rstudio-education/datascience-box) course. -->
<!-- * Exercise \@ref(ex-fish) is inspired by the Fisheries application [exercise](https://github.com/rstudio-education/datascience-box/blob/master/course-materials/application-exercises/ae-05-fisheries-datajoins/fisheries.Rmd) at the [data science in a box](https://github.com/rstudio-education/datascience-box) course. -->




I would like to thank all for their inspiration. 

<!-- Also thanks to Solveig for proofreading the draft. -->

This work is licensed under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International [CC BY-NC-SA 4.0](https://creativecommons.org/licenses/by-nc-sa/4.0/).


```{r links, child="links.md"}
```
