---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, code = readLines("setup.R"), cache = FALSE, include=FALSE}
```

```{r}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE)
```

# An introduction to Python {#sec-python}

This module gives an introduction to Python and some of the packages we use during the course.


## Learning outcomes {#sec-rl-intro-lo}

By the end of this module, you are expected to:

* Identify useful Python workflows.
* Do different programming workflows.
* Code a Python class.
* Transform data.
* Plot data.

<!-- SOLO increasing: identify · memorise · name · do simple procedure · collect data · -->
<!-- enumerate · describe · interpret · formulate · list · paraphrase · combine · do -->
<!-- algorithms · compare · contrast · explain causes · analyse · relate · derive · -->
<!-- evaluate · apply · argue · theorise · generalise · hypothesise · solve · reflect -->


## Textbook readings 

No readings and slides.


## Python and packages

An tutorial is given in this [Colab][colab-04-python] notebook. Open it and get an overview before class. 
      

# Debugging 

An tutorial is given in this [Colab][colab-04-debug-python] notebook. Open it and get an overview before class. 
      




```{r links, child="links.md", include=FALSE}
```
