---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, code = readLines("setup.R"), cache = FALSE, include=FALSE}
```

```{r, child = "_algorithms.qmd", echo=FALSE}
```

```{r}
knitr::opts_chunk$set(
   echo = TRUE
)
```

# Policy Gradient Methods {#sec-policy-gradient}

???

## Learning outcomes

After studying this chapter, you should be able to:

- 

<!-- SOLO increasing: identify · memorise · name · do simple procedure · collect data · -->

<!-- enumerate · describe · interpret · formulate · list · paraphrase · combine · do -->

<!-- algorithms · compare · contrast · explain causes · analyse · relate · derive · -->

<!-- evaluate · apply · argue · theorise · generalise · hypothesise · solve · reflect -->




## Textbook readings

For this module, you will need to read Chapter 13-13.7 in @Sutton18. Read it before continuing this module. A summary of the book notation can be seen [here][sutton-notation].

```{r, echo=FALSE}
link_slide_file_text("14", "policy-gradient")
```



## Summary 

Read Chapter 13.8 in @Sutton18.


## Exercises

`r strExercises`

You may solve the exercises in the corresponding sections in this [Colab notebook][colab-14-policy-gradient].


### Exercise

???


```{r links, child="links.md", include=FALSE}
```


