```{r, code = readLines("setup.R"), cache = FALSE, include=FALSE}
```


# Getting help {#sec-help}

We all get stuck sometimes and need some help. Below are some advises on how to help yourself and ask for help:

If you have a question:

  * Ask it during the lecture or in the breaks
  * Use generative AI (Gemini, ChatGPT, Copilot, ...) as your mentor. GAI is in general good at giving hints for programming tasks. Use it as a mentor and not to give you the solution. For instance, ask "Given me an example on a Hello world procedure in Python. Explain careful the code since I am new and want to learn." 
  * [Google is your friend](http://xkcd.com/627/). This is always the next step. Try searches like "python dfply mask", "python pandas", etc.
  * About programming errors. First try to understand the error message and solve the problem. Then you may
    * Try to debug your code by inserting `print` statements in your code or run single lines using cmd+shift+enter. 
    * Use a debugger. Further details about debugging can be seen [here][colab-04-debug-python].  
  * If you can't find an answer then it is time to ask on-line. I recommend asking a question at [stackoverflow](http://stackoverflow.com). To make your question effective, the idea is to make things as easy as possible for someone to answer. The process of providing a good minimal reproducible example (reprex) often causes you to answer your own question! See also [Stack Exchange's 'How to ask'](https://codereview.stackexchange.com/help/how-to-ask).

__Note: help using mail correspondence is not supported!__








```{r links, child="links.md"}
```
