```{r, code = readLines("setup.R"), cache = FALSE, include=FALSE}
```


# Working in groups {#groups}

[Edit]

During the course you have been allocated into groups. You are expected to solve the exercises and write the project report in your group. Before you start, it is a good idea to agree on a set of group rules. First, agree on a [coding convention](#coding-convention) when you are going to use R. Most people in the R community use snake case but camel case is also okay. Next, setup rules on when to meet and how you will organize the work. For instance, it is a good idea that all try to solve some of the exercises before you meet and you then discuss the answers, problems etc. Finally, it is a good idea to have a common place for your code. You have different options:

1) Use a cloud storage services such as Dropbox, OneDrive or Google Drive.
2) Use a version control system such as Git together with [GitHub]. GitHub is a code sharing and publishing service and may be seen as a social networking site for programmers. 
3) If you use RStudio Cloud then one person in the group can create a shared workspace with projects: 
   - First create a new workspace named e.g. _Shared_. 
   - Press _Members_ and add the group members as moderators.
   - Now go back to _Projects_ in the _RL_ workspace and move one project to the shared workspace. Rename it to e.g. _Group Project_. 
   Members will now have access to this project where you can share code. NOTE you can not work collectively on a file simultaneously. That is, only one member can change a file at a time! Hence it is a good idea to have your own private project to work on and use this project as a place where you can share code. If you want to download a project to your laptop then press the export button. 

The benefit of a cloud storage service is that it is well known to you and easy to setup. Cons are that you cannot work on the same file simultaneously. The benefit of Git and GitHub is that it manages the evolution of a set of files – called a repository – in a sane, highly structured way. If you have no idea what I'm talking about, think of it as the "Track Changes" features from Microsoft Word on steroids. Here you can work on files simultaneously. Moreover, it can be used from within RStudio. Cons are that it is harder to setup and learn. For a detailed description see [Why Git? Why GitHub?][hg-why].








# Annotate the course notes {#annotate}

I recommend using [hypothes.is] to annotate the online course notes. You can create both private and public annotations. Collaborative annotation helps people connect to each other and what they’re reading, even when they’re keeping their distance. You may also use public notes to help me indicate spell errors, unclear content etc. in the notes. 





<!-- # Schedule {}  -->

<!-- The schedule for the course is given below. -->

<!-- ```{r, echo=FALSE, message=FALSE, cache=FALSE, eval=FALSE} -->
<!-- source("common.R") -->
<!-- library(googlesheets4) -->
<!-- sheets_deauth() -->
<!-- sheet <- read_sheet("https://docs.google.com/spreadsheets/d/1-fSG3KWip0Vm5Uwf9T6vUMuH0ReBhpITHT3rl896vwc/edit?usp=sharing", -->
<!--            sheet = "Schedule") -->
<!-- knitr::kable(sheet) -->
<!-- ``` -->

<!-- <p><iframe width="1300" height="600" frameborder="0" src="https://docs.google.com/spreadsheets/d/e/2PACX-1vRka2NKE77BvtIoJc1GW-Dr_oABzQdQhMC94KpPjW7DpSTnepN9RV3m6w1ILKe3YoQwUj0C6JAe3K1T/pubhtml?gid=1591856300&amp;single=true&amp;widget=true&amp;headers=false"></iframe></p> -->




# Colophon {}

These notes was written in [bookdown](http://bookdown.org/) inside [RStudio](http://www.rstudio.com/ide/). 
<!-- The website [stat545.com](https://stat545.com) is hosted with [Netlify](https://www.netlify.com), and automatically updated after every commit by [Travis-CI](https://travis-ci.org). The complete source is available from [GitHub](https://github.com/rstudio-education/stat545). -->

This version of the notes was built with:

```{r message = FALSE, warning = FALSE, echo = FALSE}
# needed because new_session is set to true in _bookdown.yml
all_pkgs <- renv::dependencies(path = "DESCRIPTION") %>% 
  pull(Package) 

session <- devtools::session_info(pkgs = all_pkgs)

session$platform
```

Along with these packages:

```{r message = FALSE, warning = FALSE, echo = FALSE}
session$packages %>% 
  select(package, loadedversion, date, source) %>% 
  DT::datatable(rownames = FALSE,
                class = 'cell-border stripe',
                filter = list(position = 'top'),
                options = list(pageLength = nrow(session$packages), 
                           autoWidth = TRUE,
                           bInfo = FALSE,
                           paging = FALSE))
```


```{r include = FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

```{r links, child="links.md"}
```
